% the acknowledgments section

\newthought{First and foremost,} I would like to express my deepest gratitude to my adviser, Anatoly Spitkovsky, for tirelessly guiding the ``steam of my consciousness'', for trusting my intuition and allowing me to make mistakes and learn from them, for giving me freedom to do things my way, and, most importantly, for helping me keep the childlike curiosity brightly lit inside me. I am also deeply indebted to Sasha Philippov, who has patiently taught me research ethics, has shown me that asking the right questions is the crucial part of physics, and has provided me with unwavering mentorship and support throughout the years. 

I am also grateful to my thesis committee members, Jeremy Goodman, Matt Kunz, and Jim Stone, for their timely guidance and invaluable advice when I needed it most. Progress in my research would not have been possible without immense contribution from my collaborators, including Maria Petropoulou, Lorenzo Sironi, Jens Mahlmann, and many others mentioned at the beginning of each chapter of this thesis. 

It was an incredible honor and pleasure being a part of perhaps the most welcoming and bright Astro Department. I very much appreciate the positive learning environment facilitated by the faculty members, researchers and staff of this department. I would especially like to highlight my gratitude to Adam Burrows, Bruce Draine, Jenny Greene, and Michael Strauss for the incredibly interesting courses they taught. Special thanks to Polly Strauss, and Susan Duncan for patiently assisting me on so many occasions, and Steve Oakley for teaching me the proper way to handle a guitar and his unmatched musical taste. I had a great pleasure learning and developing professionally alongside my incredibly talented and bright fellow graduate students; thank you to all of you, this has been quite a journey!

I cannot even begin to describe my thanks to my dearest friends: Lev Arzamassky, for his prompt guidance and competitive spirit; Alisa Galishnikova, Kirill Lezhnin, Yoshi, and Belyash, for their heartwarming comradery during my brightest and darkest moments; Kirill Rudov, Daria Rudova, and Nina Kirillovna, for their pervasive joy that transcended my existence for the past years. I am also grateful to two people with whom I share our everlasting long-distance friendship: Ashot Grigoryan, for our long soulful discussions and conversations that unveiled for me the life-behind-mathematics; Armen Andreasyan, for our heartily enterprises and gambles that are still far from being over. I am also forever indebted to my extended family, my parents, Karine and Levon Hakobyan, and my brother, Tigran Hakobyan, for their perpetual love and support, and for being the role models I have always looked up to. 

Finally, I would like to express my deepest appreciation to my family: my wonderful cats, Pulsar and Tiri, and my wife, Arina Udalova, for her sacrifices and bravery, for taking this insane leap of faith with me, and for teaching me the most important lesson in my life -- how to be a decent human being.

